---
title: "Candy Power Ranking"
author: "Thomas Alcock"
date: "19 2 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(GGally)
library(plotly)
library(tibble)
library(tidyr)
library(gridExtra)
library(knitr)
library(forcats)
library(purrr)

library(glmnet)
library(tidymodels)
library(parsnip)
library(recipes)
library(tune)
library(rsample)
library(tune)
library(dials)
library(vip)
library(forcats)
library(gt)

```

# EDA

```{r prev}
candy_df <- read.csv(
  file = "candy-data.csv",
  header = TRUE
) %>% as_tibble()

tab_plot <- candy_df %>%
  rename(Produkt = competitorname,
         Beliebtheit = winpercent,
         Preisperzentil = pricepercent) %>% 
  mutate(Produkt = gsub("Õ","'",Produkt),
         Preisperzentil = round(100*Preisperzentil, 2),
         Beliebtheit = round(Beliebtheit, 2)) %>% 
  select(Produkt, Beliebtheit, Preisperzentil) %>% 
  arrange(desc(Beliebtheit)) %>%
  slice(1:10) %>%
  gt() %>% 
  tab_header("Top 10 Produkte nach Beliebtheit")


```

```{r eda2}
cdf_long <- pivot_longer(candy_df, cols = c(-competitorname, -sugarpercent, -winpercent, -pricepercent))

sc_plots <- function(df, cond_var, p_size = 3, opacity = 4){
  df <- df %>% 
    filter(name == cond_var) %>% 
    select(-name) %>% 
    mutate(value = ifelse(value==1, "Ja", "Nein")) %>% 
    rename(!!sym(cond_var) := value) %>% 
    mutate(!!sym(cond_var) := factor(!!sym(cond_var)))
  
  p1 <- ggplot(df, aes(x = pricepercent, y = winpercent, color = !!sym(cond_var))) +
    geom_point(size = p_size, alpha = 1/opacity) + theme_minimal() + 
    #geom_smooth(method = "lm", se = FALSE, size=0.5) +
    labs(x = "Preisperzentil", y = "Beliebtheit")
  
  p2 <- ggplot(df, aes(x = sugarpercent, y = winpercent, color = !!sym(cond_var))) +
    geom_point(size = p_size, alpha = 1/opacity) + theme_minimal() +
    #geom_smooth(method = "lm", se = FALSE, size=0.5) +
    labs(x = "Zuckergehalt", y = "Beliebtheit")
  
  grid.arrange(p1, p2)
}
```

## Boxplot

```{r}
bp <- cdf_long %>% 
  mutate(value = fct_reorder(paste0(value), winpercent),
         value = ifelse(value == 1, "Ja", "Nein")) %>% 
  ggplot(aes(x = value, y = winpercent, color = name)) + 
  geom_boxplot() +
  facet_wrap(.~name, ncol = 5) +
  theme_minimal() +
  labs(x = "Besitzt Eigenschaft", y = "Beliebtheit",  color = "")
```

```{r}
bp2 <- cdf_long %>% 
  filter(name %in% c("chocolate", "crispedricewafer", "peanutyalmondy", "fruity")) %>% 
  mutate(value = fct_reorder(paste0(value), winpercent),
         value = ifelse(value == 1, "Ja", "Nein")) %>% 
  ggplot(aes(x = value, y = winpercent, color = name)) + 
  geom_boxplot() +
  facet_wrap(.~name, ncol = 5) +
  theme_minimal() +
  labs(x = "Besitzt Eigenschaft", y = "Beliebtheit",  color = "")

bp2
```

For candies that have chocolate, rice wafers, taste fruity or are peanutyalymondy there seems to be a larger difference in winpercent.

## Scatterplots

Scatterplots for selected variables with large differences in terms of winpercent. Chocolate candies tend to be more expensive and more sugary.

```{r}
candy_df %>% 
  mutate_if(is.integer, function(x) ifelse(x==1, "Ja", "Nein")) %>% 
  pivot_longer(cols = -c(competitorname, winpercent, pricepercent, sugarpercent)) %>% 
  ggplot(aes(x = pricepercent, y = winpercent, color = value, group= name)) +
  geom_point() +
  facet_wrap(name ~ .)
```

```{r}
sc1 <- sc_plots(cdf_long, "chocolate")
```

Only few rice wafer candies are available at all, which is why we need to resample the data for estimation. Within rice wafer candies there seems to be no clear effect of price or sugar

```{r}
sc2 <- sc_plots(cdf_long, "crispedricewafer")
```

With respect to sugar and price there also seems to be no clear effect on the popularity of peanutyalmondy candies.

```{r}
sc3 <- sc_plots(cdf_long, "peanutyalmondy")
```

Fruity candies tend to be less expensive and also less popular.

```{r}
sc4 <- sc_plots(cdf_long, "fruity")
```

Gibt es Eigenschaften, die eine Imbalance aufweisen? Wichtig für Re-sampling der Daten.

```{r}
percent_yes <- candy_df %>% 
  select_if(is.integer) %>% 
  purrr::map_dfr(~ 100 * mean(.)) %>% 
  pivot_longer(cols = everything()) %>% 
  arrange(value) %>%
  ggplot(aes(x = value, y = fct_reorder(name, value))) +
  geom_bar(stat = "identity", alpha = 0.7, fill = "blue") + theme_minimal() +
  labs(x = "Relativer Anteil", y = "Eigenschaft")

percent_yes
```

# Modeling

## Bootstrap Estimation

Small sample size makes inference difficult and standard errors may be unreliable. Therefore we use a non-parametric bootstrap procedure to gauge the variation in the sampling distribution of the model coefficients.

```{r lm}
candy_df <- candy_df %>% 
  mutate(strat_var = paste0(
    hard, peanutyalmondy, caramel, 
    nougat, crispedricewafer
  ))

boots <- bootstraps(candy_df, times = 1000, strata = strat_var)

run_boot_model <- function(boot_split){
  df <- boot_split %>% analysis() %>% 
    select(-strat_var)
  simple_lm <- lm(winpercent ~ ., data = df %>% select(-competitorname))
  raw_summary <- simple_lm %>% summary()
  tidy_summary <- tidy(raw_summary)
  tidy_summary$adj_r_sq <- raw_summary$adj.r.squared
  tidy_summary$id <- boot_split$id$id
  tidy_summary
}

bs_res <- map_dfr(boots$splits, run_boot_model)
```

```{r}
plot_data <- bs_res %>% 
  filter(term != "(Intercept)")
  
bounds <- plot_data %>% 
  group_by(term) %>% 
  summarise(
    upper_bound = quantile(estimate, probs = 0.975),
    lower_bound = quantile(estimate, probs = 0.025)
  )

bs_plot <- plot_data %>% 
  select(term, estimate) %>% 
  pivot_longer(cols = estimate) %>% 
  ggplot(aes(x = value, fill = term, color = term)) + 
  geom_density(alpha = 0.7) + 
  facet_wrap(. ~ term, nrow = 4) +
  theme_minimal() + theme(legend.position = "none") +
  labs(x = "Regressionskoeffizient", y = "Dichte", title = "Koeffizienten nach Bootstrap Sampling") +
  geom_vline(data = bounds, mapping = aes(xintercept = upper_bound, color = term), linetype = 2) +
  geom_vline(data = bounds, mapping = aes(xintercept = lower_bound, color = term), linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5)
```

```{r}
bs_plot
```

```{r}
# TODO price coef lines?

bs_res %>% 
  filter(term %in% c("pricepercent", "(Intercept)"))
```

## Full model estimation

Estimate baseline model to compare against other models in robustness checks

```{r}
# basic model without bs
full_model <- lm(winpercent ~ ., 
                 data = candy_df %>% select(-competitorname, -strat_var))

full_model_res <- full_model %>% 
  summary() %>% tidy()

model_df <- candy_df %>% 
  mutate(cd = cooks.distance(full_model))

model_df <- model_df %>% 
  predict(full_model, .) %>% 
  as_tibble() %>% 
  rename(preds_baseline = value) %>% 
  bind_cols(model_df)

outlier_plot <- model_df %>% 
  ggplot(aes(x = cd)) + 
  geom_histogram(bins = 20, fill = "blue", alpha = 0.7) +
  theme_minimal() +
  labs(x = "Cook's Distanz", y = "Frequenz")
  
het_df <- model_df %>% 
  mutate(resid = winpercent - preds_baseline) %>% 
  ggplot(aes(x = preds_baseline, y = resid)) +
  geom_point(alpha = 0.7, size = 3) + theme_minimal() +
  labs(x = "Predicted", y = "Residual")
  
pred_plot <- model_df %>% 
  ggplot(aes(x = preds_baseline, y = winpercent)) +
  geom_point(size = 3, alpha = 0.5) +
  theme_minimal() +
  geom_smooth(se = FALSE, method = "lm") +
  labs(x = "Predicted", y = "Beobachtet", 
       title = paste0(
         "Adj. R-Squared:", 
          round(glance(full_model)$adj.r.squared, 2))
       )

```

## Model Summary

```{r}
coef_df <-  full_model_res %>% 
  mutate(lower = estimate - 2*std.error,
         upper = estimate + 2*std.error,
         term  = ifelse(term=="(Intercept)", "intercept", term))

coefplot <- coef_df %>%
  filter(term != "intercept") %>% 
  ggplot(aes(x = estimate, y = fct_reorder(term, abs(statistic)
  ))) + 
  geom_point(size = 4, alpha = 0.5) +
  geom_errorbar(aes(xmin = lower, xmax = upper)) +
  geom_vline(aes(xintercept = 0), linetype = 2) + 
  theme_minimal() +
  labs(x = "Effekt", y = "Attribut")
```

## Robustness checks

Do effects remain when variable composition changes slightly?

```{r}
rdf <- candy_df %>% select(-competitorname, -strat_var)

model_combs <- crossing(
  "caramel" = 0:1,
  "crispedricewafer" = 0:1,
  "nougat" = 0:1,
  "pluribus" = 0:1,
  "hard" = 0:1,
  "bar" = 0:1,
  "sugarpercent" = 0:1,
  "pricepercent" = 0:1,
  "peanutyalmondy" = 0:1,
  "fruity" = 0:1,
  "chocolate" = 0:1
)

# exclude degenerate cases, i.e. all 1 or all 0
model_combs <- model_combs[-c(1, nrow(model_combs)),]

drop_vars <- function(idx){
  model_combs[idx,] %>% 
  pivot_longer(cols = everything()) %>% 
  filter(value == 1) %>% 
  pull(name)
}

rb_lm <- function(x){
  drop_these <- drop_vars(x)
  df <- rdf %>% select(-all_of(drop_these))
  res <- tidy(summary(lm(winpercent ~ ., data = df)))
  res$missing_vars <- paste0(drop_these, collapse = "_")
  res
}

robust_models <- map_dfr(1:nrow(model_combs), rb_lm)
```

Distributions of coefficients taking out combinations of variables and then estimating a model without them.

```{r}
rob_plot <- robust_models %>%
  filter(term != "(Intercept)") %>% 
  ggplot(aes(x = estimate, y = ..density.., fill = term)) + 
  geom_density(aes(y = ..density.., color = term), alpha = 0.4) +
  theme_minimal() +
  labs(x = "Effekt", y = "Frequenz") +
  facet_wrap(. ~ term, nrow = 5) +
  geom_vline(aes(xintercept = 0), linetype = 2) +
  theme(legend.position = "none")
```

The effect for fruity flavor seems to vary a great deal across different model specifications, same is true for price, bar, rice wafers and nougat.

**chocolate peanutyalmondy** and **sugarpercent** seem to have pretty robust effects in terms of sign, with chocolate displaying some more variation in terms of effect size.

## Interaction Model

Interaction terms have large standard errors, no conclusion can be drawn abt interaction terms fruit x price, fruit x sugar.

```{r}
int_df <- candy_df %>% select(-competitorname, -strat_var) 

int_df <- int_df %>% mutate(
  fruit_price = pricepercent * fruity,
  fruit_sugar = sugarpercent * fruity,
  choc_price = chocolate * pricepercent,
  choch_sugar = chocolate * sugarpercent
)

int_model <- lm(winpercent ~ ., data = int_df)

int_model_res <- int_model %>% 
  summary() %>% 
  tidy() %>% 
  mutate(
    upper = estimate + 2*std.error,
    lower = estimate - 2*std.error
  )

broom::glance(int_model)
```

# Bootstrap vs Full model Coefs

```{r}
mean_df <- bs_res %>% 
  group_by(term) %>% 
  summarise(mean_estimate = mean(estimate)) %>% 
  ungroup() %>% 
  left_join(bounds) %>% 
  left_join(coef_df) %>% 
  filter(term != "(Intercept)")
    

bs_regular_comparison <- mean_df %>% 
  ggplot(data = mean_df, mapping = aes(y = fct_reorder(term, estimate))) +
  geom_point(mapping = aes(x = estimate, color = "Standard OLS"), size = 4, alpha = 0.4) +
  geom_point(mapping = aes(x = mean_estimate, color = "Bootstrapping"), size = 4, alpha = 0.4) +
  geom_errorbar(aes(xmin = lower_bound, xmax = upper_bound), color = "blue") + 
  geom_errorbar(aes(xmin = lower, xmax = upper), color = "red") +
  theme_minimal() + labs(x = "Effekt", y = "Attribut", color = "")  +
  geom_vline(aes(xintercept = 0), linetype= 2) +
  scale_x_continuous(breaks  = pretty_breaks(10)) +
  scale_color_manual(
    values = c("Standard OLS" = "red", "Bootstrapping" = "blue")
  )

```

# Recommendation

```{r}
model_df %>% 
  filter(
    chocolate == 1 & 
    peanutyalmondy == 1 &
    hard == 0
  ) %>% 
  arrange(desc(sugarpercent)) %>% 
  slice(1) %>% 
  t()
```

```{r}
sim_df <- crossing(
  "caramel" = 0:1,
  "crispedricewafer" = 0:1,
  "nougat" = 0:1,
  "pluribus" = 0:1,
  "hard" = 0:1,
  "bar" = 0:1,
  "pricepercent" = seq(min(candy_df$pricepercent), max(candy_df$pricepercent), 0.01),
  "sugarpercent" = seq(min(candy_df$sugarpercent), max(candy_df$sugarpercent), 0.01),
  "peanutyalmondy" = 0:1,
  "fruity" = 0:1,
  "chocolate" = 0:1
)

sim_df$pred <- predict.lm(full_model, newdata = sim_df)


sim_df %>% 
  filter(pred == max(pred)) %>% 
  pivot_longer(cols = everything()) %>% 
  filter(name %in% c("chocolate", "peanutyalmondy", "sugarpercent", ""))
  knitr::kable()
```

```{r}
model_df %>% 
  filter(chocolate == 1 &
         peanutyalmondy == 1 &
         hard == 0 &
         nougat == 01) %>% 
  arrange(pricepercent)
```
